Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-23 papers/FX/draft.tex
Reshuffled the sections in the FX paper.

@@

%   where
%   \[
%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
%   \]
%   Now notice
%   \[
%   \ln(y e^{-y}) = \ln(y) - y
%   \]
%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
%   function. Hence we have
%   \begin{eqnarray*}
%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
%   \end{eqnarray*}
%   Now it is clear
%   \[
%   \nu\exp\left\{
%   W\left[
%     -\left(1 + {1 \over \nu}\right)
%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
%   \right]
%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
%   &=& W(-y e^{-y}) + y
% \end{eqnarray*}
\section{Review of multivariate GARCH models:
  same tail index or not}
Many multivariate GARCH models imply a common tail index for all
series under consideration, e.g. the Orthogonal GARCH model of
Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
well as the Full Factor GARCH model of Vrontos et al (2003)
\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
models are characterized by their treatment of each series as a linear
combination of factors, and each of the factors is modeled as a GARCH
process. As a result of this treatment, each and every series under
consideration necessarily has the lowest of the tail indices of its
contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
as Andersen, Davis, Kreiss and Mikosch (2009)
\cite{andersen2009handbook}).

\section{Data analysis}
\subsection{Hill Estimator}
\label{sec:Hill}
There have been several estimators for the tail index proposed in the
literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
arguably Hill (1975) \cite{hill1975simple}:
\[
\hat \alpha = \left[
  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
  \right]^{-1}
\]
where $x_{(i)}$ is the $i$-th upper order statistic of a sample
$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
applicable to 
\begin{itemize}
\item iid random variables (Mason (1982) \cite{mason1982laws})
\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
  and de Haan \cite{rootzen1992tail})
\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
  \cite{resnick1995consistency}).
\end{itemize}
In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
\[
\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
\]

@@

Figure \ref{fig:Energy_alpha_K},
\ref{fig:Consumer_Staples_alpha_K},
and \ref{fig:Information_Technology_alpha_K}
show the estimates of $K$ alongside the Hill estimates of $\alpha$ of
the ``energy'', ``consumer staples'' and ``information
technology'' sectors of the S\&P 500 index, respectively.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
  \caption{Hybrid estimates of the lower-tail parameters of daily
    return series in the {\it Energy} sector of S\&P
    500 index.
  }
  \label{fig:Energy_alpha_K}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
  \caption{Hybrid estimates of the lower-tail parameters of daily
    return series in the {\it Consumer Staples} sector of S\&P
    500 index.
  }
  \label{fig:Consumer_Staples_alpha_K}
\end{figure}

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
  \caption{Hybrid estimates of the lower-tail parameters of daily
    return series in the {\it Information Technology} sector of S\&P
    500 index.
  }
  \label{fig:Information_Technology_alpha_K}
\end{figure}

@@

\section{Conclusion}
TODO: Summarize and conclude that all correlated series share the same
tail index; their tail functions differ from each other via their
scale parameters.

>>>>

\section{Empirical Results}
\subsection{Hill estimates of lower tail indices}
\label{sec:Hill}
There have been several estimators for the tail index proposed in the
literature, e.g. Hill (1975) \cite{hill1975simple}, Pikands (1975)
\cite{pickands1975statistical}, Dekkers and Einmahl and De Haan (1989)
\cite{dekkers1989moment}. See also Embrechts and Kl\"uppelberg and
Mikosch (1997)\cite{Embrechts1997}. The most popular of these is
arguably Hill (1975) \cite{hill1975simple}:
\[
\hat \alpha = \left[
  {1 \over k} \sum_{i=1}^k \ln {x_{(i)} \over x_{(k+1)}}
  \right]^{-1}
\]
where $x_{(i)}$ is the $i$-th upper order statistic of a sample
$x_1, x_2, ..., x_n$. It has been shown that the Hill estimator is
applicable to 
\begin{itemize}
\item iid random variables (Mason (1982) \cite{mason1982laws})
\item a weakly dependent sequence (Rootz{\'e}n, Leadbetter,
  and de Haan \cite{rootzen1992tail})
\item a linear process (Resnick and St{\u{a}}ric{\u{a}} (1995)
  \cite{resnick1995consistency}).
\end{itemize}
In these cases it is shown that, if $k \to \infty$ and $k/n \to 0$
as $n \to \infty$, the Hill estimator is weakly consistent, i.e.
\[
\sqrt k (\hat \alpha - \alpha) \overset{d}{\to} N(0, \alpha^2)
\]

\section{Review of multivariate GARCH models:
  same tail index or not}
Many multivariate GARCH models imply a common tail index for all
series under consideration, e.g. the Orthogonal GARCH model of
Alexander and Chibumba (1997) \cite{alexander1997multivariate}, its
generalization GO-GARCH of van der Weide (2002) \cite{van2002go}, as
well as the Full Factor GARCH model of Vrontos et al (2003)
\cite{vrontos2003full} and the Generalized Orthogonal Factor GARCH
model of Lanne and Saikkonen (2007) \cite{lanne2007modeling}. These
models are characterized by their treatment of each series as a linear
combination of factors, and each of the factors is modeled as a GARCH
process. As a result of this treatment, each and every series under
consideration necessarily has the lowest of the tail indices of its
contributing factors (cf. Mikosch (2013) \cite{Mikosch2013} as well
as Andersen, Davis, Kreiss and Mikosch (2009)
\cite{andersen2009handbook}).

@@

@@

Figure \ref{fig:Energy_alpha_K},
\ref{fig:Consumer_Staples_alpha_K},
and \ref{fig:Information_Technology_alpha_K}
show the estimates of $K$ alongside the Hill estimates of $\alpha$ of
the ``energy'', ``consumer staples'' and ``information
technology'' sectors of the S\&P 500 index, respectively.
\begin{figure}[htb!]
  \centering
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_alpha_K.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_alpha_K}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Energy_alpha_K_ci.pdf}
                    \caption{``Energy''}
                    \label{fig:Energy_alpha_K_ci}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_alpha_K.pdf}
                    \caption{``Consumer Staples''}
                    \label{fig:Consumer_Staples_alpha_K}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Consumer_Staples_alpha_K_ci.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Consumer_Staples_alpha_K_ci}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_alpha_K.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Information_Technology_alpha_K}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\linewidth}
    \includegraphics[width=\textwidth]
                    {Information_Technology_alpha_K_ci.pdf}
                    \caption{``Information Technology''}
                    \label{fig:Information_Technology_alpha_K_ci}
  \end{subfigure}
  \caption{\small Maximum likelihood estimates of $K$, the scale
    parameter, of stocks in sectors of the S\&P 500 index. The points
    are the
    estimated values; the red bars are the upper bound of the 98\%
    confidence interval. The lower bound is 0.
  }
\end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Energy_alpha_K.pdf}
%%   \caption{Hybrid estimates of the lower-tail parameters of daily
%%     return series in the {\it Energy} sector of S\&P
%%     500 index.
%%   }
%%   \label{fig:Energy_alpha_K}
%% \end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Consumer_Staples_alpha_K.pdf}
%%   \caption{Hybrid estimates of the lower-tail parameters of daily
%%     return series in the {\it Consumer Staples} sector of S\&P
%%     500 index.
%%   }
%%   \label{fig:Consumer_Staples_alpha_K}
%% \end{figure}

%% \begin{figure}[htb!]
%%   \centering
%%   \includegraphics[width=\textwidth]{Information_Technology_alpha_K.pdf}
%%   \caption{Hybrid estimates of the lower-tail parameters of daily
%%     return series in the {\it Information Technology} sector of S\&P
%%     500 index.
%%   }
%%   \label{fig:Information_Technology_alpha_K}
%% \end{figure}

@@

%   where
%   \[
%   y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
%   \]
%   Now notice
%   \[
%   \ln(y e^{-y}) = \ln(y) - y
%   \]
%   is a decreasing function. Thus $-y e^{-y}$ is an increasing
%   function. Hence we have
%   \begin{eqnarray*}
%     W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
%   \end{eqnarray*}
%   Now it is clear
%   \[
%   \nu\exp\left\{
%   W\left[
%     -\left(1 + {1 \over \nu}\right)
%     e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
%   \right]
%   + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
%   &=& W(-y e^{-y}) + y
% \end{eqnarray*}
\section{Conclusion}
TODO: Summarize and conclude that all correlated series share the same
tail index; their tail functions differ from each other via their
scale parameters.
